---
title: "NC ERP | Nonverbal category knowledge limits the amount of information encoded in object representations"
autohr: Barbara Pomiechowska, Teodora Gliga
output: 
 html_document:
    toc: yes
    toc_float: yes
---

```{r message=FALSE, warning=FALSE, , paged.print=FALSE, include=FALSE}
library(readxl)
library(lsr) 
library(tidyr)
library(plyr)
library(ggpubr)
library(ggplot2)
library(emmeans) 
library(afex)
library(rstatix)
library(Rmisc) 
library(tidyverse)
library(car) 
library(sjstats)
```

# Data

Read in data.
```{r include=FALSE}
#Set directory
wd <- getwd()
setwd(wd)

#Read in data
dfnc  <-  read_excel("data/data_erp.xlsx", sheet = "nc")
```

Check data structure.
```{r}
dfnc$Experiment <- factor(dfnc$Experiment)
dfnc$Category <- factor(dfnc$Category)
str(dfnc)
```


## Data structure

The file *datasummary.xlsx* contains the average Nc amplitude data in the sheet *nc*:

* $ID$                  -- unique ID identifying each participant

* $Experiment$          -- Experiment 1, Experiment 2

* $Category$            -- Familiar, Unfamiliar

* $NC changeNO$        -- average Nc amplitudes in the no change condition

* $NC changeWITHIN$    -- average Nc amplitudes in the within-category change condition

* $NC changeACROSS$    -- average Nc amplitudes in the across-category change condition


# Experiment 1

```{r}
longe1 <- dfnc %>%
  gather(key = OcclusionOutcome, value = nc, c("NC_changeNO","NC_changeWITHIN","NC_changeACROSS")) %>%
  filter(Experiment == "Experiment 1")
```

## Normality

```{r echo=FALSE}
longe1 %>%
  group_by(OcclusionOutcome, Category) %>%
  shapiro_test(nc)
```

## Omnibus ANOVA

- $OcclusionOutcome$ (no change v. within-category change v. across-category change) as a within-subject factor
- $Category$ (familiar v. unfamiliar) as a between-subject factor

```{r warning=FALSE}
#Run the analysis
ae1 <- aov_car(nc ~ OcclusionOutcome*Category + 
               Error(ID/OcclusionOutcome) + Category, 
             data = longe1)
#Display the results
summary(ae1)
eta_sq(ae1, partial = TRUE)
```

## Familiar categories

### ANOVA

```{r}
longe1familiar <- filter(longe1, Category == "Familiar")
```

- $OcclusionOutcome$ (no change v. within-category change v. across-category change) as a within-subject factor

```{r warning=FALSE}
#Run the analysis
ae1familiar <- aov_car(nc ~ OcclusionOutcome + Error(ID/OcclusionOutcome), 
             data = longe1familiar)
#Display the results
summary(ae1familiar)
eta_sq(ae1familiar, partial = TRUE)
```

#### Followup t tests

```{r}
e1familiar <- dfnc %>%
  filter(Experiment == "Experiment 1") %>%
  filter(Category == "Familiar")
```

No change v. across-categry change:
```{r}
t.test(e1familiar$NC_changeNO,e1familiar$NC_changeACROSS, paired = TRUE)
cohensD(e1familiar$NC_changeNO,e1familiar$NC_changeACROSS, method = "paired")
```

No change v. within-categry change:
```{r}
t.test(e1familiar$NC_changeNO,e1familiar$NC_changeWITHIN, paired = TRUE)
cohensD(e1familiar$NC_changeNO,e1familiar$NC_changeWITHIN, method = "paired")
```

Across-category change v. within-categry change:
```{r}
t.test(e1familiar$NC_changeWITHIN,e1familiar$NC_changeACROSS, paired = TRUE)
cohensD(e1familiar$NC_changeWITHIN,e1familiar$NC_changeACROSS, method = "paired")
```

## Unfamiliar categories

### ANOVA

```{r}
longe1unfamiliar <- filter(longe1, Category == "Unfamiliar")
```

- $OcclusionOutcome$ (no change v. within-category change v. across-category change) as a within-subject factor

```{r warning=FALSE}
#Run the analysis
ae1unfamiliar <- aov_car(nc ~ OcclusionOutcome + Error(ID/OcclusionOutcome), 
             data = longe1unfamiliar)
#Display the results
summary(ae1unfamiliar)
eta_sq(ae1unfamiliar, partial = TRUE)
```

#### Followup t tests

```{r}
e1unfamiliar <- dfnc %>%
  filter(Experiment == "Experiment 1") %>%
  filter(Category == "Unfamiliar")
```

No change v. across-categry change:
```{r}
t.test(e1unfamiliar$NC_changeNO,e1unfamiliar$NC_changeACROSS, paired = TRUE)
```

No change v. within-categry change:
```{r}
t.test(e1unfamiliar$NC_changeNO,e1unfamiliar$NC_changeWITHIN, paired = TRUE)
```

Across-category change v. within-categry change:
```{r}
t.test(e1unfamiliar$NC_changeWITHIN,e1unfamiliar$NC_changeACROSS, paired = TRUE)
```

# Experiment 2

```{r}
longe2 <- dfnc %>%
  gather(key = OcclusionOutcome, value = nc, c("NC_changeNO","NC_changeWITHIN","NC_changeACROSS")) %>%
  filter(Experiment == "Experiment 2")
```

## Normality

```{r echo=FALSE}
longe2 %>%
  group_by(OcclusionOutcome, Category) %>%
  shapiro_test(nc)
```

## Omnibus ANOVA

- $OcclusionOutcome$ (no change v. within-category change v. across-category change) as a within-subject factor
- $Category$ (familiar v. unfamiliar) as a between-subject factor

```{r warning=FALSE}
#Run the analysis
ae2 <- aov_car(nc ~ OcclusionOutcome*Category + 
               Error(ID/OcclusionOutcome) + Category, 
             data = longe2)
#Display the results
summary(ae2)
eta_sq(ae2, partial = TRUE)
```

## Familiar categories / Category-training

### ANOVA

```{r}
longe2familiar <- filter(longe2, Category == "Familiar")
```

- $OcclusionOutcome$ (no change v. within-category change v. across-category change) as a within-subject factor

```{r warning=FALSE}
#Run the analysis
ae2familiar <- aov_car(nc ~ OcclusionOutcome + Error(ID/OcclusionOutcome), 
             data = longe2familiar)
#Display the results
summary(ae2familiar)
eta_sq(ae2familiar, partial = TRUE)
```

#### Followup t tests

```{r}
e2familiar <- dfnc %>%
  filter(Experiment == "Experiment 2") %>%
  filter(Category == "Familiar")
```

No change v. across-categry change:
```{r}
t.test(e2familiar$NC_changeNO,e2familiar$NC_changeACROSS, paired = TRUE)
```

No change v. within-categry change:
```{r}
t.test(e2familiar$NC_changeNO,e2familiar$NC_changeWITHIN, paired = TRUE)
```

Across-category change v. within-categry change:
```{r}
t.test(e2familiar$NC_changeWITHIN,e2familiar$NC_changeACROSS, paired = TRUE)
```

## Unfamiliar categories / Control condition

### ANOVA

```{r}
e2unfamiliar <- filter(longe2, Category == "Unfamiliar")
```

- $OcclusionOutcome$ (no change v. within-category change v. across-category change) as a within-subject factor

```{r warning=FALSE}
#Run the analysis
ae2unfamiliar <- aov_car(nc ~ OcclusionOutcome + Error(ID/OcclusionOutcome), 
             data = e2unfamiliar)
#Display the results
summary(ae2unfamiliar)
eta_sq(ae2unfamiliar, partial = TRUE)
```

#### Followup t tests

```{r}
e2unfamiliar <- dfnc %>%
  filter(Experiment == "Experiment 2") %>%
  filter(Category == "Unfamiliar")
```

No change v. across-categry change:
```{r}
t.test(e2unfamiliar$NC_changeNO,e2unfamiliar$NC_changeACROSS, paired = TRUE)
```

No change v. within-categry change:
```{r}
t.test(e2unfamiliar$NC_changeNO,e2unfamiliar$NC_changeWITHIN, paired = TRUE)
```

Across-category change v. within-categry change:
```{r}
t.test(e2unfamiliar$NC_changeWITHIN,e2unfamiliar$NC_changeACROSS, paired = TRUE)
```

# Experiments 1-2 | comparison

```{r}
long <- dfnc %>%
  gather(key = OcclusionOutcome, value = nc, c("NC_changeNO","NC_changeWITHIN","NC_changeACROSS"))
```

## ANOVA

- $OcclusionOutcome$ (no change v. within-category change v. across-category change) as a within-subject factor
- $Category$ (familiar v. unfamiliar) as a between-subject factor
- $Experiment$ (1 v.2) as a between-subject factor

```{r warning=FALSE}
#Run the analysis
ae1e2 <- aov_car(nc ~ OcclusionOutcome*Category*Experiment + 
               Error(ID/OcclusionOutcome) + Category + Experiment, 
             data = long)
#Display the results
summary(ae1e2)
eta_sq(ae1e2, partial = TRUE)
```

### Followup tests

#### Familiar categories

```{r}
familiar <- dfnc %>%
  filter(Category == "Familiar")
```

No change v. across-categry change:
```{r}
t.test(familiar$NC_changeNO,familiar$NC_changeACROSS, paired = TRUE)
cohensD(familiar$NC_changeNO,familiar$NC_changeACROSS, method = "paired")
```

No change v. within-categry change:
```{r}
t.test(familiar$NC_changeNO,familiar$NC_changeWITHIN, paired = TRUE)
```

Across-category change v. within-categry change:
```{r}
t.test(familiar$NC_changeWITHIN,familiar$NC_changeACROSS, paired = TRUE)
```

#### Unfamiliar categories

```{r}
unfamiliar <- dfnc %>%
  filter(Category == "Unfamiliar")
```

No change v. across-categry change:
```{r}
t.test(unfamiliar$NC_changeNO,unfamiliar$NC_changeACROSS, paired = TRUE)
cohensD(unfamiliar$NC_changeNO,unfamiliar$NC_changeACROSS, method = "paired")
```

No change v. within-categry change:
```{r}
t.test(unfamiliar$NC_changeNO,unfamiliar$NC_changeWITHIN, paired = TRUE)
```

Across-category change v. within-categry change:
```{r}
t.test(unfamiliar$NC_changeWITHIN,unfamiliar$NC_changeACROSS, paired = TRUE)
```

