---
title: "Time-freqeuncy data | Nonverbal category knowledge limits the amount of information encoded in object representation"
autohr: Barbara Pomiechowska, Teodora Gliga
output: 
 html_document:
    toc: yes
    toc_float: yes
---

```{r}
library(readxl)
library(tidyverse)
library(ggpubr)
library(rstatix)
library(afex)
library(sjstats)
library(car)
library(lsr) 
```

# Data structure

The files *data_timefreq_gamma.xlsx*, *data_timefreq_alpha.xlsx* contain the average time-frequency data in the gamma (25-45 Hz) and alpha (6-9 Hz) bands. These data are stored in the *summary* sheet and have the following structure:

* $ID$                  -- unique ID identifying each participant

* $Experiment$          -- Experiment 1, Experiment 2

* $Category$            -- Familiar, Unfamiliar

* $Presentation\_left$   -- average activation in the left aoi during presentation (object in full view after sliding onto the stage)

* $Presentation\_right$  -- average activation in the right aoi during presentation (object in full view after sliding onto the stage)

* $Occlusion\_left$      -- average activation in the left aoi during occlusion (object fully invisible after the occluder went up)

* $Occlusion\_right$     -- average activation in the right aoi during occlusion (object fully invisible after the occluder went up)

* $Trials$              -- number of artifact-free trials contributed by each infant

The remaining sheets (*E1_familiar*, *E1_unfamiliar*, *E2_familiar*, *E2_unfamiliar*) contain individual channel data (averaged for each of the respective frequency ranges).

# Read in the data

Read in data (create two data frames: dfgamma, dfalpha).
```{r include=FALSE}
#Set directory
wd <- getwd()
setwd(wd)

#Read in data
dfgamma  <-  read_excel("data/data_timefreq_gamma.xlsx", sheet = "summary")
dfalpha  <-  read_excel("data/data_timefreq_alpha.xlsx", sheet = "summary")
```

Review data structure: dfgamma.
```{r}
dfgamma$Experiment <- factor(dfgamma$Experiment)
dfgamma$Category <- factor(dfgamma$Category)
str(dfgamma)
```

Review data structure: dfalpha
```{r}
dfalpha$Experiment <- factor(dfalpha$Experiment)
dfalpha$Category <- factor(dfalpha$Category)
str(dfgamma)
```

# Gamma (25-45 Hz)

```{r}
longgamma <- dfgamma %>%
  gather(key = Condition, value = gamma, c("Presentation_left","Presentation_right",
                                            "Occlusion_left","Occlusion_right")) %>%
  mutate(Phase=ifelse(str_detect(Condition,"Presentation"),"Presentation","Occlusion")) %>%
  mutate(Hemisphere = ifelse(str_detect(Condition,"left"),"Left","Right"))
```

## Omnibus ANOVA

```{r}
#Normality
longgamma %>%
  group_by(Experiment, Phase, Category, Hemisphere, Condition) %>%
  shapiro_test(gamma)
```

- $Experiment$ (1 v. 2) as a between-subject factor
- $Category$ (familiar v. unfamiliar) as a between-subject factor
- $Hemisphere$ (left v. right) as a within-subject factor
- $Phase$ (presentation v. occlusion) as a within-subject factor

```{r warning=FALSE}
#Run the analysis
aogamma <- aov_car(gamma ~ Experiment*Category*Hemisphere*Phase + 
               Error(ID/Hemisphere*Phase) + Experiment + Category, 
             data = longgamma)
#Display the results
summary(aogamma)
eta_sq(aogamma, partial = TRUE)
```

## Split by Phase

### Presentation

Compute mean activation across hemispheres.
Save it in a new variable (MEAN_presentation).
```{r}
dfgamma <- dfgamma %>%
  mutate(MEAN_presentation = rowMeans(dfgamma[4:5]))
```

#### ANOVA

```{r}
leveneTest(MEAN_presentation ~ Experiment*Category, data = dfgamma)
```

Run the ANOVA.

- $Experiment$ (1 v. 2) as a between-subject factor
- $Category$ (familiar v. unfamiliar) as a between-subject factor

```{r}
#Run the analysis
agammapres <- aov_car(MEAN_presentation ~ Experiment*Category + 
               Error(ID) + Experiment + Category, 
             data = dfgamma)
#Display the results
summary(agammapres)
eta_sq(agammapres, partial = TRUE)
```

#### Comparisons to baseline

```{r}
dfgammafamiliar <- filter(dfgamma, Category == "Familiar")
t.test(dfgammafamiliar$MEAN_presentation, mu = 0, alternative = "two.sided")
cohensD(dfgammafamiliar$MEAN_presentation, mu = 0)
```

```{r}
dfgammaunfamiliar <- filter(dfgamma, Category == "Unfamiliar")
t.test(dfgammaunfamiliar$MEAN_presentation, mu = 0, alternative = "two.sided")
cohensD(dfgammaunfamiliar$MEAN_presentation, mu = 0)
```

#### Figures

Plot.
```{r}
plotgamma <- dfgamma %>%
  group_by(Category) %>%
  get_summary_stats(MEAN_presentation,type = "mean_se")

pgamma <- ggplot(data = plotgamma,aes(x=Category, y = mean, color=Category, fill=Category,  group=Category)) +
   theme_light() +
          theme(strip.text = element_text(size = 12),
                panel.grid.major.y = element_line(colour="grey75"),
                panel.grid.minor.y = element_blank(),
                panel.grid.major.x = element_blank(),
                axis.text.x= element_text(size=14, angle=0),
                axis.text.y= element_text(size=14),
                axis.title.y = element_text(size = 14, family = "Helvetica"),
                legend.position="none",
                axis.title.x = element_blank()) +
  geom_bar(stat="identity", position=position_dodge(),width=.7,alpha =.2) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.1,position=position_dodge(.7)) +
  scale_fill_manual(values = c("black","gray60", "black", "gray60")) +
  scale_color_manual(values = c("black","gray60", "black", "gray60")) +
  scale_y_continuous(name=(gamma~"-band activity("~mu~"V)"), limits=c(-.1, .2)) 

pgamma
```

```{r}
m <- stat_summary(geom ="point", fun.y=mean, shape=15, size=4, color="yellow", fill="red", 
                  position = position_dodge(width = 0.75)) 

boxgamma <- ggplot(data = dfgamma, aes(x=Category, y=MEAN_presentation)) +
  theme_light() +
          theme(strip.text = element_text(size = 12),
                panel.grid.major.y = element_line(colour="grey75"),
                panel.grid.minor.y = element_blank(),
                panel.grid.major.x = element_blank(),
                axis.text.x= element_text(size=14, angle=0),
                axis.text.y= element_text(size=14),
                axis.title.y = element_text(size = 14, family = "Helvetica"),
                legend.position="none",
                axis.title.x = element_blank()) +
  geom_boxplot(aes(fill=Category), position = position_dodge(.9), 
               alpha = .2, outlier.shape = NA, width = .5) +
  geom_jitter(aes(color=Category),position = position_jitterdodge(.15), size =1) +
  scale_y_continuous(name=gamma~"-band activity("~mu~"V)", limits=c(-.4, .4)) +
  scale_fill_manual(values = c("black", "gray60")) +
  scale_color_manual(values = c("black", "gray60")) 

boxgamma
```

```{r echo=FALSE}
fgp <- ggarrange(pgamma, boxgamma,
          ncol = 2, widths = c(1.5,1),
          labels = c("1", "2"),
          common.legend = TRUE, legend = "none")

annotate_figure(fgp,
               top = text_grob("Presentation: object fully visible", color = "black", face = "bold", size = 16))

```

```{r include=FALSE}
ggsave("plots_timefreq/fig3gammapresentation.png", plot = last_plot(), width = 17.5, height = 10, units = c("cm"))
```


### Occlusion

Compute mean activation across hemispheres.
Save it in a new variable (MEAN_occlusion).

```{r}
dfgamma <- dfgamma %>%
  mutate(MEAN_occlusion = rowMeans(dfgamma[6:7]))
```

#### ANOVA

```{r}
leveneTest(MEAN_occlusion ~ Experiment*Category, data = dfgamma)
```

Run the ANOVA.

- $Experiment$ (1 v. 2) as a between-subject factor
- $Category$ (familiar v. unfamiliar) as a between-subject factor

```{r}
#Run the analysis
agammaoccl <- aov_car(MEAN_occlusion ~ Experiment*Category + 
               Error(ID) + Experiment + Category, 
             data = dfgamma)
#Display the results
summary(agammaoccl)
eta_sq(agammaoccl, partial = TRUE)
```

#### Figures

Plot.
```{r}
plotgammaocclusion <- dfgamma %>%
  group_by(Category) %>%
  get_summary_stats(MEAN_occlusion,type = "mean_se")

pgammaocclusion <- ggplot(data = plotgammaocclusion,aes(x=Category, y = mean, color=Category, fill=Category,  group=Category)) +
   theme_light() +
          theme(strip.text = element_text(size = 12),
                panel.grid.major.y = element_line(colour="grey75"),
                panel.grid.minor.y = element_blank(),
                panel.grid.major.x = element_blank(),
                axis.text.x= element_text(size=12, angle=0),
                axis.text.y= element_text(size=14),
                axis.title.y = element_text(size = 14, family = "Helvetica"),
                legend.position="none",
                axis.title.x = element_blank()) +
  geom_bar(stat="identity", position=position_dodge(),width=.7,alpha =.2) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.1,position=position_dodge(.7)) +
  scale_fill_manual(values = c("black","gray60", "black")) +
  scale_color_manual(values = c("black","gray60", "black")) +
  scale_y_continuous(name="Aamplitude (Hz)", limits=c(-.1, .2)) 

pgammaocclusion
```

```{r}
m <- stat_summary(geom ="point", fun.y=mean, shape=23, size=4, color="black", fill="yellow", 
                  position = position_dodge(width = 0.75)) 

boxgammaocclusion <- ggplot(data = dfgamma, aes(x=Category, y=MEAN_occlusion)) +
  theme_light() +
          theme(strip.text = element_text(size = 12),
                panel.grid.major.y = element_line(colour="grey75"),
                panel.grid.minor.y = element_blank(),
                panel.grid.major.x = element_blank(),
                axis.text.x= element_text(size=12.5, angle=0),
                axis.text.y= element_text(size=14),
                axis.title.y = element_text(size = 14, family = "Helvetica"),
                legend.position="none",
                axis.title.x = element_blank()) +
  geom_boxplot(aes(fill=Category), position = position_dodge(.9), 
               alpha = .2, outlier.shape = NA, width = .5) +
  geom_jitter(aes(color=Category),position = position_jitterdodge(.25), size = 1) +
  scale_y_continuous(name="Amplitude (Hz)", limits=c(-.5, .5)) +
  scale_fill_manual(values = c("black", "gray60")) +
  scale_color_manual(values = c("black","gray60")) 

boxgammaocclusion
```

```{r echo=FALSE}
fgp <- ggarrange(pgammaocclusion, boxgammaocclusion,
          ncol = 2, widths = c(1.5,1),
          labels = c("1", "2"),
          common.legend = TRUE, legend = "none")

annotate_figure(fgp,
               top = text_grob("Occlusion: object fully invisible", color = "black", face = "bold", size = 16))

```

```{r include=FALSE}
ggsave("plots_timefreq/figu3gammaocclusion.png", plot = last_plot(), width = 17.5, height = 10, units = c("cm"))
```


# Alpha (6-9 Hz)

## Omnibus ANOVA

```{r}
longalpha <- dfalpha %>%
  gather(key = Condition, value = alpha, c("Presentation_left","Presentation_right",
                                            "Occlusion_left","Occlusion_right")) %>%
  mutate(Phase=ifelse(str_detect(Condition,"Presentation"),"Presentaiton","Occlusion")) %>%
  mutate(Hemisphere = ifelse(str_detect(Condition,"left"),"Left","Right"))
```

```{r}
longalpha %>%
  group_by(Experiment, Phase, Category, Hemisphere, Condition) %>%
  shapiro_test(alpha)
```

- $Experiment$ (1 v. 2) as a between-subject factor
- $Category$ (familiar v. unfamiliar) as a between-subject factor
- $Hemisphere$ (left v. right) as a within-subject factor
- $Phase$ (presentation v. occlusion) as a within-subject factor

```{r}
#Run the analysis
aoalpha <- aov_car(alpha ~ Experiment*Category*Hemisphere*Phase + 
               Error(ID/Hemisphere*Phase) + Experiment + Category, 
             data = longalpha)
#Display the results
summary(aoalpha)
eta_sq(aoalpha, partial = TRUE)
```

## Split by Phas

### Presentation

Compute mean activation across hemispheres.
Save it in a new variable (MEAN_presentation).
```{r}
dfalpha <- dfalpha %>%
  mutate(MEAN_presentation = rowMeans(dfalpha[4:5]))
```

```{r}
longalphapres <- dfalpha %>%
  gather(key = Hemisphere, value = alpha, c("Presentation_left","Presentation_right"))
```

#### ANOVA

- $Experiment$ (1 v. 2) as a between-subject factor
- $Category$ (familiar v. unfamiliar) as a between-subject factor
- $Hemisphere$ (left v. right) as a within-subject factor

```{r warning=FALSE}
#Run the analysis
aalphapres <- aov_car(alpha ~ Experiment*Category*Hemisphere + 
               Error(ID/Hemisphere) + Experiment + Category, 
             data = longalphapres)
#Display the results
summary(aalphapres)
```

#### Comparisons to baseline

```{r}
t.test(dfalpha$MEAN_presentation, mu = 0, alternative = "two.sided")
cohensD(dfalpha$MEAN_presentation, mu = 0)
```

#### Figures

Plot.
```{r}
plotalphapresentation <- dfalpha %>%
  group_by(Category) %>%
  get_summary_stats(MEAN_presentation,type = "mean_se")

palphapresentation <- ggplot(data = plotalphapresentation,
                          aes(x=Category, y = mean, color=Category, fill=Category,  group=Category)) +
   theme_light() +
          theme(strip.text = element_text(size = 12),
                panel.grid.major.y = element_line(colour="grey75"),
                panel.grid.minor.y = element_blank(),
                panel.grid.major.x = element_blank(),
                axis.text.x= element_text(size=12, angle=0),
                axis.text.y= element_text(size=14),
                axis.title.y = element_text(size = 14, family = "Helvetica"),
                legend.position="bottom",
                axis.title.x = element_blank()) +
  geom_bar(stat="identity", position=position_dodge(),width=.7,alpha =.2) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.1,position=position_dodge(.7)) +
  scale_fill_manual(values = c("black","gray60", "black","gray60")) +
  scale_color_manual(values = c("black","gray60", "black","gray60")) +
  scale_y_continuous(name=alpha~"-band activity("~mu~"V)", limits=c(-.6, .6)) 

palphapresentation
```


```{r}
m <- stat_summary(geom ="point", fun.y=mean, shape=23, size=4, color="black", fill="yellow", 
                  position = position_dodge(width = 0.75)) 

boxalphapres <- ggplot(data = dfalpha,
                            aes(x=Category, y=MEAN_presentation, fill = Category, color = Category)) +
  theme_light() +
          theme(strip.text = element_text(size = 12),
                panel.grid.major.y = element_line(colour="grey75"),
                panel.grid.minor.y = element_blank(),
                panel.grid.major.x = element_blank(),
                axis.text.x= element_text(size=12.5, angle=0),
                axis.text.y= element_text(size=14),
                axis.title.y = element_text(size = 14, family = "Helvetica"),
                legend.position="bottom",
                axis.title.x = element_blank()) +
  geom_boxplot(aes(fill=Category), position = position_dodge(.9), 
               alpha = .2, outlier.shape = NA, width = .5) +
  geom_jitter(aes(color=Category),position = position_jitterdodge(.25), size = 1) +
  scale_y_continuous(name=alpha~"-band activity("~mu~"V)", limits=c(-2, 2)) +
  scale_fill_manual(values = c("black", "gray60","black", "gray60")) +
  scale_color_manual(values = c("black","gray60","black", "gray60")) 

boxalphapres
```


```{r echo=FALSE}
fgp <- ggarrange(palphapresentation, boxalphapres,
          ncol = 2, widths = c(1.5,1),
          labels = c("1", "2"),
          common.legend = TRUE, legend = "bottom")

annotate_figure(fgp,
               top = text_grob("Presentation: object fully visible", color = "black", face = "bold", size = 16))

```

```{r include=FALSE}
ggsave("plots_timefreq/fig3alphapresentation.png", plot = last_plot(), width = 17.5, height = 10, units = c("cm"))
```


### Occlusion

```{r}
dfalpha <- dfalpha %>%
  mutate(MEAN_occlusion = rowMeans(dfalpha[6:7]))
```

```{r}
longalphaocclusion <- dfalpha %>%
  gather(key = Hemisphere, value = occlusion_alpha, c("Occlusion_left","Occlusion_right"))

longalphaocclusion$Hemisphere <- factor(longalphaocclusion$Hemisphere, levels = c("Occlusion_left","Occlusion_right"),
                                        labels = c("Left", "Right"))
```

#### ANOVA

- $Experiment$ (1 v. 2) as a between-subject factor
- $Category$ (familiar v. unfamiliar) as a between-subject factor
- $Hemisphere$ (left v. right) as a within-subject factor

```{r warning=FALSE}
#Run the analysis
aalphaoccl <- aov_car(occlusion_alpha ~ Experiment*Category*Hemisphere + 
               Error(ID/Hemisphere) + Experiment + Category, 
             data = longalphaocclusion)
#Display the results
summary(aalphaoccl)
eta_sq(aalphaoccl, partial = TRUE)
```

#### Figures

```{r}
plotalphaocclusion <- dfalpha %>%
  group_by(Category) %>%
  get_summary_stats(MEAN_occlusion,type = "mean_se")

palphaocclusion <- ggplot(data = plotalphaocclusion,
                          aes(x=Category, y = mean, color=Category, fill=Category,  group=Category)) +
   theme_light() +
          theme(strip.text = element_text(size = 12),
                panel.grid.major.y = element_line(colour="grey75"),
                panel.grid.minor.y = element_blank(),
                panel.grid.major.x = element_blank(),
                axis.text.x= element_text(size=12, angle=0),
                axis.text.y= element_text(size=14),
                axis.title.y = element_text(size = 14, family = "Helvetica"),
                legend.position="bottom",
                axis.title.x = element_blank()) +
  geom_bar(stat="identity", position=position_dodge(),width=.7,alpha =.2) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.1,position=position_dodge(.7)) +
  scale_fill_manual(values = c("black","gray60", "black","gray60")) +
  scale_color_manual(values = c("black","gray60", "black","gray60")) +
  scale_y_continuous(name=alpha~"-band activity("~mu~"V)", limits=c(-.6, .6)) 

palphaocclusion
```


```{r}
m <- stat_summary(geom ="point", fun.y=mean, shape=23, size=4, color="black", fill="yellow", 
                  position = position_dodge(width = 0.75)) 

boxalphaocclusion <- ggplot(data = dfalpha,
                            aes(x=Category, y=MEAN_occlusion, fill = Category, color = Category)) +
  theme_light() +
          theme(strip.text = element_text(size = 12),
                panel.grid.major.y = element_line(colour="grey75"),
                panel.grid.minor.y = element_blank(),
                panel.grid.major.x = element_blank(),
                axis.text.x= element_text(size=12.5, angle=0),
                axis.text.y= element_text(size=14),
                axis.title.y = element_text(size = 14, family = "Helvetica"),
                legend.position="bottom",
                axis.title.x = element_blank()) +
  geom_boxplot(aes(fill=Category), position = position_dodge(.9), 
               alpha = .2, outlier.shape = NA, width = .5) +
  geom_jitter(aes(color=Category),position = position_jitterdodge(.25), size = 1) +
  scale_y_continuous(name=alpha~"-band activity("~mu~"V)", limits=c(-2, 2)) +
  scale_fill_manual(values = c("black", "gray60","black", "gray60")) +
  scale_color_manual(values = c("black","gray60","black", "gray60")) 

boxalphaocclusion
```

```{r echo=FALSE}
fgp <- ggarrange(palphaocclusion, boxalphaocclusion,
          ncol = 2, widths = c(1.5,1),
          labels = c("1", "2"),
          common.legend = TRUE, legend = "bottom")

annotate_figure(fgp,
               top = text_grob("Occlusion: object fully invisible", color = "black", face = "bold", size = 16))

```

```{r include=FALSE}
ggsave("plots_timefreq/fig3alphaocclusion.png", plot = last_plot(), width = 17.5, height = 10, units = c("cm"))
```

#### Comparisons to baseline


```{r}
aux <- dfalpha %>% filter(Category == "Familiar")
t.test(aux$MEAN_occlusion, mu = 0, alternative = "two.sided")
```

```{r}
aux <- dfalpha %>% filter(Category == "Unfamiliar")
t.test(aux$MEAN_occlusion, mu = 0, alternative = "two.sided")
cohensD(aux$MEAN_occlusion, mu = 0)
```

```{r}
dfalphafamiliar <- subset(dfalpha, Category == "Familiar")
dfalphaunfamiliar <- subset(dfalpha, Category == "Unfamiliar")
```

```{r}
t.test(dfalphafamiliar$Occlusion_left, mu = 0, alternative = "two.sided")
t.test(dfalphafamiliar$Occlusion_right, mu = 0, alternative = "two.sided")
```

```{r}
t.test(dfalphaunfamiliar$Occlusion_left, mu = 0, alternative = "two.sided")
cohensD(dfalphaunfamiliar$Occlusion_left, mu = 0)
t.test(dfalphaunfamiliar$Occlusion_right, mu = 0, alternative = "two.sided")
cohensD(dfalphaunfamiliar$Occlusion_right, mu = 0)
```
